Low-Thrust Trajectory Design Near Asteroids Using Reachability Sets

### 2016-08-05 manuscript.tex
Adding more detail on dynamics and polyhedron model

@@

The gravitational attraction due to a constant density polyhedron is given as
\begin{align*}
	U(\vecbf{r}) &= 1\frac{1}{2} G \sigma \sum_{e \in \text{edges}} \vecbf{r}_e \cdot \vecbf{E}_e \cdot \vecbf{r}_e \cdot L_e - \frac{1}{2}G \sigma \sum_{f \in \text{faces}} \vecbf{r}_f \cdot \vecbf{F}_f \cdot \vecbf{r}_f \cdot \omega_f 
\end{align*}
where \( \vecbf{r}_e, \vecbf{r}_f \) is the vector from the spacecraft to any point on the respective edge or face.
The polyhedron is defined by an array of vectors in the body fixed frame which define the vertices of the body.
Triangular faces are defined as the plane created by any three vertices. 
Consider three vectors \( \vecbf{v}_0, \vecbf{v}_1, \vecbf{v}_2 \), assumed to be ordered in a counterclockwise direction, which define a face. 
The outward normal vector to the face and to each edge is defined as
\begin{align*}
	\hat{\vecbf{n}}_f &= \parenth{\vecbf{v}_{1} - \vecbf{v}_0} \times \parenth{\vecbf{v}_{2} - \vecbf{v}_1} ,\\
	\hat{\vecbf{n}}_{e} &= \parenth{\vecbf{v}_{i+1} - \vecbf{v}_i} \times \hat{\vecbf{n}}_f .
\end{align*}
Using these normal vectors we define two dyads
\begin{align*}
	\vecbf{E}_e &= \hat{\vecbf{n}}_A \hat{\vecbf{n}}_{12}^A + \hat{\vecbf{n}}_B \hat{\vecbf{n}}_{21}^B ,\\
	\vecbf{F}_f &= \hat{\vecbf{n}}_f \hat{\vecbf{n}}_f .
\end{align*}
The edge dyad \( \vecbf{E}_e \) is defined for each edge and a function of the two adjacent faces meeting at the edge. 
The face dyad \( \vecbf{F}_f \) is defined for each face and is a function of the face normal vector.

Let \( \vecbf{r}_i \) be the vector from the spacecraft to the vertex \( \vecbf{v}_i \) and it's length is \( r_i = \norm{\vecbf{r}_i} \).
The per-edge factor \( L_e \), for the edge connecting vertices \( \vecbf{v}_i \) and \( \vecbf{v}_j \), with a constant length \( e_{ij} \) is 
\begin{align*}
	L_e &= \ln \frac{r_i + r_j + e_{ij}}{r_i + r_j - e_{ij}}.
\end{align*}

For the face defined by the vertices \( \vecbf{v}_i, \vecbf{v}_j, \vecbf{v}_k \) the per-face factor \( \omega_f \) is
\begin{align*}
	\omega_f &= 2 \arctan \frac{\vecbf{r}_i \cdot \vecbf{r}_j \times \vecbf{r}_k}{r_i r_j r_k + r_i \parenth{\vecbf{r}_j \cdot \vecbf{r}_k} + r_j \parenth{\vecbf{r}_k \cdot \vecbf{r}_i} + r_k \parenth{\vecbf{r}_i \cdot \vecbf{r}_j}} .
\end{align*}

Using these definitions, the attraction, gravity gradient matrix, and Laplacian are defined as
\begin{align*}
	\nabla U ( \vecbf{r} ) &= -G \sigma \sum_{e \in \text{edges}} \vecbf{E}_e \cdot \vecbf{r}_e \cdot L_e + G \sigma \sum_{f \in \text{faces}} \vecbf{F}_f \cdot \vecbf{r}_f \cdot \omega_f ,\\
	\nabla \nabla U ( \vecbf{r} ) &= G \sigma \sum_{e \in \text{edges}} \vecbf{E}_e  \cdot L_e - G \sigma \sum_{f \in \text{faces}} \vecbf{F}_f \cdot \omega_f , \\
	\nabla^2 U &= -G \sigma \sum_{f \in \text{faces}}  \omega_f .
\end{align*}

One interesting thing to note is that both \( \vecbf{E}, \vecbf{F} \) can be precomputed without knowledge of the position of the satellite. 
They are both solely functions of the vertices and edges of the polyhedral shape model.
One a position vector \( \vecbf{r} \) is defined then the scalars \( \omega_f, L_e \) can be computed for each face and edge.

Duplicate the equations for polyhedron model

Give an outline of how to compute the potential (which equations and order)

Cite the original werner papers

@@

We define a reference frame originating at the center of mass of the asteroid.
The body-fixed reference frame is composed of the unit vectors \( \hat{\vecbf{x}} , \hat{\vecbf{y}}, \hat{\vecbf{z}} \), which are aligned along the principal axes of smallest, intermediate, and largest moment of inertia, respectively.
The body-fixed equations of motion of a massless particle about an arbitrarily rotating asteroid are given by
\begin{align}\label{eq:body_eoms}
	\ddot{\vecbf{r}} + 2 \vecbf{\Omega} \times \dot{\vecbf{r}} + \vecbf{\Omega} \times \parenth{ \vecbf{\Omega} \times \vecbf{r} } + \dot{\vecbf{\Omega}} \times \vecbf{r} = U_{\vecbf{r}} , 
\end{align}
where \( \vecbf{\Omega} \) is the instaneous angular velocity vector of the asteroid represented in the body-fixed frame, \( \vecbf{r} \) is the position of the particle in the body-fixed frame, and \( U_{\vecbf{r}} \) is the gradient of the gravitational potential~\cite{greenwood1988}.
We assume that the asteroid rotates at a uniform rate, \( \norm{\vecbf{\Omega}} = \omega \), about the axis of the maximum moment of inertia, \( \vecbf{\Omega} = \omega \hat{\vecbf{z} }\).
As a result, we can represent the equations of motion in scalar form as
\begin{align} \label{eq:eoms}
	\begin{split}
    	\ddot{x} - 2 \omega \dot{y} - \omega^2 y &= U_x , \\
    	\ddot{y} + 2 \omega \dot{x} - \omega^2 x &= U_y , \\
    	\ddot{z} &= U_z .
	\end{split}
\end{align}

We assume that in our problem we are modelling a spacecraft rather than a massless particle. 
In this situation, the state is defined as \( \vecbf{x} = \begin{bmatrix}\vecbf{r} &\vecbf{v} \end{bmatrix}^T\) with \(\vecbf{r} = \begin{bmatrix} x & y & z \end{bmatrix}^T \in \R^{3\times1}\) and \(\vecbf{v}= \begin{bmatrix} \dot{x} & \dot{y} & \dot{z} \end{bmatrix}^T \in \R^{3\times1}\) representing the position and velocity with respect to the body-fixed frame, respectively.
We further assume that our spacecraft is capable of exerting a translational force, \( \vecbf{u} \in \R^{3\times1} \), in any direction, while subject to a maximum magnitude constraint.
This is typical of many spacecraft which offer full rotational freedom and can direct a maximum, and potentially varying, thrust in any direction.
The equations of motion may be rewritten in state space form as
\begin{align}\label{eq:state_space_eoms}
	\begin{bmatrix} \dot{\vecbf{r}} \\ \dot{\vecbf{v}} \end{bmatrix} &=
\end{align}
where the matrix \( G \) represents the

Cite scheeres and his book

@@

Define \Poincare section

\subsection{Optimal Control Formulation}

>>>>

Mention we're looking at asteroid 4769 Castalia

@@

We represent the gravitational potential of the asteroid using a polyhedron gravitation model.
This model is composed of a polyhedron, a three-dimensional solid body, which is defined by a series of vectors in the body-fixed frame.
The vectors define vertices in the body-fixed frame as well as planar faces which compose the surface of the asteroid.
We assume that each face is a triangle composed of three vertices and three edges.
As a result, only two faces meet at each edge while three faces meet at each vertex. 
Several asteroids have been accurately mapped from a combination of Earth based sensors as well as visiting spacecraft (CITE \href{http://sbn.psi.edu/pds/resource/rshape.html}{http://sbn.psi.edu/pds/resource/rshape.html}). 
Only the body-fixed vectors, and their associated topology, is required to define the exterior gravitational model.
References~\citenum{werner1994} and~\citenum{werner1997} give a detailed derivation of the polyhedron model from first principles.
Here, we summarize the key developments and equations required for implementation and clarity.

Insert picture of face and vectors

Consider three vectors \( \vecbf{v}_1, \vecbf{v}_2, \vecbf{v}_3 \), assumed to be ordered in a counterclockwise direction, which define a face. 
It is easy to define the three edges of each face as
\begin{align}\label{eq:edges}
	\vecbf{e}_{i+1,i} = \vecbf{v}_{i+1} - \vecbf{v}_i ,
\end{align}
where the index \( i \in \parenth{1,2,3} \) is used to permute all edges of each face. 
Since each edge is a member of two faces, there exist two edges which are defined in opposite directions between the same vertices.
We can also define the outward normal vector to face \( f\)  as
\begin{align}\label{eq:face_normal}
	\hat{\vecbf{n}}_f &= \parenth{\vecbf{v}_{2} - \vecbf{v}_1} \times \parenth{\vecbf{v}_{3} - \vecbf{v}_2} ,
\end{align}
and the outward facing normal vector to each edge as
\begin{align}\label{eq:edge_normal}
	\hat{\vecbf{n}}_{i+1,i}^f &= \parenth{\vecbf{v}_{i+1} - \vecbf{v}_i} \times \hat{\vecbf{n}}_f .
\end{align}
For each face we define the face dyad \( \vecbf{F}_f \) as
\begin{align}\label{eq:face_dyad}
	\vecbf{F}_f &= \hat{\vecbf{n}}_f \hat{\vecbf{n}}_f .
\end{align}
Each edge is a member of two faces and has an outward pointing edge normal vector, given in~\cref{eq:edge_normal}, perpindicular to both the edge and the face normal.
For the edge connecting the vectors \( \vecbf{v}_1 \) and \( \vecbf{v}_2 \) which are shared between the faces \(A\) and \( B\) the per edge dyad is given by
\begin{align}\label{eq:edge_dyad}
	\vecbf{E}_{12} = \hat{\vecbf{n}}_A \hat{\vecbf{n}}_{12}^A + \hat{\vecbf{n}}_B \hat{\vecbf{n}}_{21}^B .
\end{align}
The edge dyad \( \vecbf{E}_e \) is defined for each edge and a function of the two adjacent faces meeting at the edge. 
The face dyad \( \vecbf{F}_f \) is defined for each face and is a function of the face normal vector.

Let \( \vecbf{r}_i \) be the vector from the spacecraft to the vertex \( \vecbf{v}_i \) and it's length is \( r_i = \norm{\vecbf{r}_i} \).
The per-edge factor \( L_e \), for the edge connecting vertices \( \vecbf{v}_i \) and \( \vecbf{v}_j \), with a constant length \( e_{ij} \) is 
\begin{align}\label{eq:edge_factor}
	L_e &= \ln \frac{r_i + r_j + e_{ij}}{r_i + r_j - e_{ij}}.
\end{align}
For the face defined by the vertices \( \vecbf{v}_i, \vecbf{v}_j, \vecbf{v}_k \) the per-face factor \( \omega_f \) is
\begin{align}\label{eq:face_factor}
	\omega_f &= 2 \arctan \frac{\vecbf{r}_i \cdot \vecbf{r}_j \times \vecbf{r}_k}{r_i r_j r_k + r_i \parenth{\vecbf{r}_j \cdot \vecbf{r}_k} + r_j \parenth{\vecbf{r}_k \cdot \vecbf{r}_i} + r_k \parenth{\vecbf{r}_i \cdot \vecbf{r}_j}} .
\end{align}
The gravitational potential due to a constant density polyhedron is given as
\begin{align}\label{eq:potential}
	U(\vecbf{r}) &= 1\frac{1}{2} G \sigma \sum_{e \in \text{edges}} \vecbf{r}_e \cdot \vecbf{E}_e \cdot \vecbf{r}_e \cdot L_e - \frac{1}{2}G \sigma \sum_{f \in \text{faces}} \vecbf{r}_f \cdot \vecbf{F}_f \cdot \vecbf{r}_f \cdot \omega_f 
\end{align}
where \( \vecbf{r}_e, \vecbf{r}_f \) is the vector from the spacecraft to any point on the respective edge or face.
Furthermore we can use these definitions to define the attraction, gravity gradient matrix, and Laplacian are defined as
\begin{align}
	\nabla U ( \vecbf{r} ) &= -G \sigma \sum_{e \in \text{edges}} \vecbf{E}_e \cdot \vecbf{r}_e \cdot L_e + G \sigma \sum_{f \in \text{faces}} \vecbf{F}_f \cdot \vecbf{r}_f \cdot \omega_f , \label{eq:attraction}\\
	\nabla \nabla U ( \vecbf{r} ) &= G \sigma \sum_{e \in \text{edges}} \vecbf{E}_e  \cdot L_e - G \sigma \sum_{f \in \text{faces}} \vecbf{F}_f \cdot \omega_f , \label{eq:gradient_matrix}\\
	\nabla^2 U &= -G \sigma \sum_{f \in \text{faces}}  \omega_f \label{eq:laplacian}.
\end{align}

One interesting thing to note is that both~\cref{eq:face_dyad,eq:edge_dyad} can be precomputed without knowledge of the position of the satellite. 
They are both solely functions of the vertices and edges of the polyhedral shape model and are computed once and stored.
Once a position vector \( \vecbf{r} \) is defined then the scalars~\cref{eq:edge_factor,eq:face_factor} can be computed for each face and edge.
Finally,~\cref{eq:potential} is used to compute the gravitational potential on the spacecraft.

@@

We define a reference frame originating at the center of mass of the asteroid.
The body-fixed reference frame is composed of the unit vectors \( \hat{\vecbf{x}} , \hat{\vecbf{y}}, \hat{\vecbf{z}} \), which are aligned along the principal axes of smallest, intermediate, and largest moment of inertia, respectively.
The body-fixed equations of motion of a massless particle about an arbitrarily rotating asteroid are given by
\begin{align}\label{eq:body_eoms}
	\ddot{\vecbf{r}} + 2 \vecbf{\Omega} \times \dot{\vecbf{r}} + \vecbf{\Omega} \times \parenth{ \vecbf{\Omega} \times \vecbf{r} } + \dot{\vecbf{\Omega}} \times \vecbf{r} = U_{\vecbf{r}} , 
\end{align}
where \( \vecbf{\Omega} \) is the instaneous angular velocity vector of the asteroid represented in the body-fixed frame, \( \vecbf{r} \) is the position of the particle in the body-fixed frame, and \( U_{\vecbf{r}} \) is the gradient of the gravitational potential~\cite{scheeres2012a}.
We assume that the asteroid rotates at a uniform rate, \( \norm{\vecbf{\Omega}} = \omega \), about the axis of the maximum moment of inertia, \( \vecbf{\Omega} = \omega \hat{\vecbf{z} }\).
As a result, we can represent the equations of motion in scalar form as
\begin{align} \label{eq:eoms}
	\begin{split}
    	\ddot{x} - 2 \omega \dot{y} - \omega^2 y &= U_x , \\
    	\ddot{y} + 2 \omega \dot{x} - \omega^2 x &= U_y , \\
    	\ddot{z} &= U_z .
	\end{split}
\end{align}

We assume that in our problem we are modelling a spacecraft rather than a massless particle. 
In this situation, the state is defined as \( \vecbf{x} = \begin{bmatrix}\vecbf{r} &\vecbf{v} \end{bmatrix}^T\) with \(\vecbf{r} = \begin{bmatrix} x & y & z \end{bmatrix}^T \in \R^{3\times1}\) and \(\vecbf{v}= \begin{bmatrix} \dot{x} & \dot{y} & \dot{z} \end{bmatrix}^T \in \R^{3\times1}\) representing the position and velocity with respect to the body-fixed frame, respectively.
We further assume that our spacecraft is capable of exerting a translational force, \( \vecbf{u} \in \R^{3\times1} \), in any direction, while subject to a maximum magnitude constraint.
This is typical of many spacecraft which offer full rotational freedom and can direct a maximum, and potentially varying, thrust in any direction.
The equations of motion may be rewritten in state space form as
\begin{align}\label{eq:state_space_eoms}
	\begin{bmatrix} \dot{\vecbf{r}} \\ \dot{\vecbf{v}} \end{bmatrix} &=
	\begin{bmatrix}\vecbf{v} \\ \vecbf{g} \parenth{\vecbf{r}} + \vecbf{h}\parenth{\vecbf{v}} + \vecbf{u} \end{bmatrix} ,
\end{align}
where the terms \(\vecbf{g} \parenth{\vecbf{r}} \) and \( \vecbf{h}\parenth{\vecbf{v}} \) are given by
\begin{align}\label{eq:state_space_terms}
	\vecbf{g}\parenth{\vecbf{r}} = \begin{bmatrix}  U_x + \omega^2 x \\ U_y + \omega^2 y \\ U_z \end{bmatrix} ,\quad
	\vecbf{h}\parenth{\vecbf{r}} = \begin{bmatrix} 2 \omega \dot{y} \\ -2 \omega \dot{x} \\ 0 \end{bmatrix} .
\end{align}

Since Castalia is a uniformly rotating asteroid, the equations of motion are time invariant when represented in the body-fixed frame.
In addition, there exists an integral of motion, or a conserved quantity, the is constant for all motion of a particle, absent any other perturbing forces.
The Jacobi constant, \( J (\vecbf{r} , \vecbf{v} ) \), is given by
\begin{align}\label{eq:jacobi}
	J \parenth{\vecbf{r}, \vecbf{v}} = \frac{1}{2} \omega^2 \parenth{x^2 + y^2} + U(\vecbf{r}) - \frac{1}{2} \parenth{\dot{x}^2 + \dot{y}^2 + \dot{z}^2} .
\end{align}
The Jacobi constant functions in a similar manner as used in three-body problem~\cite{szebehely1967}.
\Cref{eq:jacobi} divides the phase space into distinct realms of possible motion.
We can define zero-velocity surfaces using the Jacobi constant by fixing the value to a desired constant.
The zero-velocity surfaces are the locus of points where the kinetic energy and hence velocity vanishes.

Figure of zero velocity surfaces for each of the three planes

@@

Define \Poincare section following the method of Scheeres

4 dimensional \Poincare section so we need to discretize the direction using three angles
